embedplyr: Tools for Working With Text Embeddings

### 2025-11-10 paper/paper.Rmd
references formatting
update acknowledgements

Existing tools for working with embeddings in R are generally specific to particular model architectures. For example, **word2vec** [@wijffels2023] provides access to word2vec models [@mikolov2013a; @mikolov2013b], **text2vec** [@selivanov2023] provides access to LDA, LSA, and GloVe models [@blei2003; @deerwester1990; @pennington2014], **fastText** [@mouselimis2024] provides access to fastText models [@facebook2016; @bojanowski2017a; @bojanowski2017b; @grave2018], and **text** [@kjell2023] provides access to transformers-based LLMs [@wolf2020]. While these tools are already invaluable in the social and behavioral sciences, using different syntax and separate data structures for each model architecture can be cumbersome. This is especially true given the prevalence of studies that make use of multiple model architectures in parallel analyses [e.g. @carrella2023; @hussain2024; @markus2024]. **embedplyr** fills this gap with a model-agnostic approach; it can be used to work with embeddings from any model framework using syntax that will be familiar to any **tidyverse** user. While some existing tools focus on convenience functions for encouraging expert-recommended best practices [e.g. @kjell2023], **embedplyr** prioritizes flexibility---much like its namesake, **dplyr** [@wickham2023]. This approach yields simple, modular functions that are useful both for educating students [see @teitelbaum2024] and experimenting with novel methods.

@@

>>>>

Existing tools for working with embeddings in R are generally specific to particular model architectures. For example, **word2vec** [@wijffels2023] provides access to word2vec models [@mikolov2013a; @mikolov2013b], **text2vec** [@selivanov2023] provides access to LDA, LSA, and GloVe models [@blei2003; @deerwester1990; @pennington2014], **fastText** [@mouselimis2024] provides access to fastText models [@facebook2016; @bojanowski2017a; @bojanowski2017b; @grave2018], and **text** [@kjell2023] provides access to transformers-based LLMs [@wolf2020]. While these tools are already invaluable in the social and behavioral sciences, using different syntax and separate data structures for each model architecture can be cumbersome. This is especially true given the prevalence of studies that make use of multiple model architectures in parallel analyses [e.g. @carrella2023; @hussain2024; @markus2024]. **embedplyr** fills this gap with a model-agnostic approach; it can be used to work with embeddings from any model framework using syntax that will be familiar to any **tidyverse** user. While some existing tools focus on convenience functions for encouraging expert-recommended best practices [e.g. @kjell2023], **embedplyr** prioritizes flexibility---much like its namesake, **dplyr** [@wickham2025]. This approach yields simple, modular functions that are useful both for educating students [see @teitelbaum2024] and experimenting with novel methods.

@@

# Acknowledgements

The author thanks Almog Simchon for helpful feedback in the early stages of development. This work was supported in part by a grant by the Ministry of Aliyah and Integration of the State of Israel.
